---
title: "trzeciak_homework3"
author: "Jagoda Trzeciak"
date: "2025-04-09"
output: html_document
---

# Homework Problem 1
Analyze the 68k PBMCs dataset in the same way as presented in the Seurat's guide with PBMC3k. Apply QC, PCA, jackstraw, clustering, and t-SNE to create figure similar to Figure 3b on Zheng et al. 2017. Note that there are differences between Zheng's original analysis and Seurat's analysis. Pay attentions to hyper-parameters that you must choose for this new bigger dataset.

Present the t-SNE visualization with 10 clusters as defined by K-means clustering

Reproduce Figure 3 but note difference in results: https://www.nature.com/articles/ncomms14049/figures/3


## Load the data
```{r}
library(Seurat)
library(dplyr)
library(patchwork)

pbmc.data <- Read10X(data.dir = "fresh_68k_pbmc_donor_a_raw_gene_bc_matrices/matrices_mex/hg19")

pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc68k", min.cells = 3, min.features = 200)
pbmc
```

## QC
```{r}
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3,pt.size=0.05,alpha=0.1)
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt") 
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") 
plot1 + plot2

pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
```

## Normalizing the data
```{r}
pbmc <- NormalizeData(pbmc)
```
## Identification of highly variable features (feature selection)
```{r}
pbmc <- FindVariableFeatures(pbmc, selection.method = 'vst', nfeatures = 2000)

top10 <- head(VariableFeatures(pbmc), 10)
plot1 <- VariableFeaturePlot(pbmc)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## Scaling the data
```{r}
all.genes <- rownames(pbmc)
# pbmc <- ScaleData(pbmc, features = all.genes)
pbmc <- ScaleData(pbmc)
```
## Perform linear dimensional reduction
```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
print(pbmc[['pca']], dims = 1:5, nfeatures = 5)
VizDimLoadings(pbmc, dims = 1:2, reduction = 'pca')
DimPlot(pbmc, reduction = 'pca')
DimHeatmap(pbmc, dims = 1, cells = 500, balanced = TRUE)
```

## Jackstraw
```{r}
pbmc <- JackStraw(pbmc, num.replicate = 20)
pbmc <- ScoreJackStraw(pbmc, dims = 1:20)
JackStrawPlot(pbmc, dims = 1:20)
ElbowPlot(pbmc)
```

## Clustering
```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.2)
```
## tSNE
```{r}
library(ggplot2)

set.seed(42)

pbmc <- RunTSNE(pbmc, reduction = 'pca', dims=1:10, reduction.name='tsne')
pca.embeddings <- Embeddings(pbmc, reduction = "pca")[, 1:20]
kmeans.res <- kmeans(pca.embeddings, centers = 10)
pbmc$kmeans_cluster <- as.factor(kmeans.res$cluster)

cluster_freq <- prop.table(table(pbmc$kmeans_cluster)) * 100
new_labels <- paste0("Cluster ", names(cluster_freq), "\n", round(cluster_freq, 1), "%")
names(new_labels) <- names(cluster_freq)

pbmc_clustering <- SetIdent(pbmc, value = "kmeans_cluster")
pbmc_clustering <- RenameIdents(pbmc_clustering, new_labels)

pdf("Trzeciak_problem1.pdf")
DimPlot(pbmc_clustering, reduction = "tsne", label = TRUE, pt.size = 0.5)  + ggtitle("Trzeciak_problem1")
dev.off()
```

# Homework problem 2
Create a hierachical clustering by applying K-means clustering to cells defined by each of 10 cluster. Try to find a suitable number of clusters (k) for each sub-population.
```{r}
library(cluster)

pbmc$sub_cluster <- NA
main_clusters <- levels(pbmc$kmeans_cluster)

pbmc <- SetIdent(pbmc, value = "kmeans_cluster")
for (cl in main_clusters) {
  cells.in.cluster <- WhichCells(pbmc, idents = cl)
  cluster.pca <- pca.embeddings[cells.in.cluster, ]
  
  best_sil <- -1
  best_k <- 2
  for (k in 2:5) {
    km <- kmeans(cluster.pca, centers = k)
    sil <- silhouette(km$cluster, dist(cluster.pca))
    
    if (mean(sil[,3]) > best_sil) {
      best_sil <- mean(sil[,3])
      best_k <- k
    }
  }
  
  set.seed(42)
  kmeans.res <- kmeans(cluster.pca, centers = best_k)
  
  pbmc$sub_cluster[cells.in.cluster] <- paste0("C", cl, "_", kmeans.res$cluster)
}
```

```{r}
library(patchwork)

pdf("Trzeciak_problem2.pdf", width = 12, height = 12)

overall_plot <- DimPlot(pbmc, reduction = "tsne", group.by = "sub_cluster", label = TRUE, repel = TRUE) +
  ggtitle("Overall t-SNE with all clusters")
print(overall_plot)

subplots <- list()
for (cl in main_clusters) {
  p <- DimPlot(pbmc, reduction = "tsne", cells = WhichCells(pbmc, idents = cl), group.by = "sub_cluster") +
      ggtitle(paste("Subclusters of Cluster", cl)) + theme(plot.title = element_text(size = 12))
  subplots[[cl]] <- p
}

print(wrap_plots(subplots, ncol = 2))

dev.off()
```